### 2025-02-15 ca.stereo.qmd
make it so

---
title: "CA of Music Genre Stereotypes"
author: "Omar Lizardo"
date: "`r Sys.Date()`"
execute: 
  eval: true
  echo: true
  output: true
  warning: false
  message: false
---

```{r setup, include=FALSE}
   library(conflicted)
   library(haven)
   library(here)
   library(dplyr)
```

```{r import data and select genre stereotype variables}
   #importing data
   dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
   #replacing ids with character names
   dat <- bind_cols(tibble(id_char = paste("id", sapply(dat[, 1], as.character), sep = "_")), dat) %>% 
   select(id_char, classicalfemale:hvymtluc, -ends_with("none")) %>%  #selecting relevant columns
   na.omit() #removing missing values
   head(dat)
```

```{r}
   make.mat <- function(x) {
      matrix(as.numeric(x), 20, 15, byrow = TRUE)
   }
   r.proj <- function(x) {
      x %*% t(x)
   }
   c.proj <- function(x) {
      t(x) %*% x
   }
   ind.list <- split(dat[, 2:301], seq(nrow(dat)))
   ind.mat <- lapply(ind.list, make.mat)
   ind.row <- lapply(ind.mat, r.proj)
   ind.col <- lapply(ind.mat, c.proj)
```