### 2025-02-15 create-match.qmd

---
title: "Jilbert's Distinction Centrality"
author: "Omar Lizardo"
date: "`r Sys.Date()`"
execute: 
  eval: true
  echo: true
  output: true
  warning: false
  message: false
---

```{r setup, include=FALSE}
   library(conflicted)
   library(haven)
   library(here)
   library(dplyr)
```
```{r import data and select genre stereotype variables}
   #importing data
   dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
   #replacing ids with character names
   dat <- bind_cols(tibble(id_char = paste("id", sapply(dat[, 1], as.character), sep = "_")), dat) %>% 
   select(id_char, classicalfemale:hvymtluc, -ends_with("none")) %>%  #selecting relevant columns
   na.omit() #removing missing values
   head(dat)
```

```{r creating match matrix}
   n <- nrow(dat)
   match <- matrix(0, n, n)
   for (i in 1:n) {
      for (j in 1:n) {
         if (i < j) {
            match[i, j] <- 300 - (sum(abs(dat[i, 2:301] - dat[j, 2:301])))
            print(i)
            print(j)
         }
      }
   }
```

```{r}
   match.f <- match
   match.f[lower.tri(match.f)] <- t(match)[lower.tri(match)]
   match.s <- (2*(match.f/300) - 1)
   save(match.f, match.s, file = "match.rda")
```