### 2025-05-03 heidelberg-slides.qmd

---
title: "Musical Genres as Fuzzy Categories"
subtitle: ""
author: "Omar Lizardo"
institute: "*University of California, Los Angeles*"
date: "`r Sys.Date()`"
format:
  revealjs:
    incremental: false 
    theme: simple 
    smaller: true
    slide-number: true
    slide-level: 2
    progress: true
    footer: "https://github.com/olizardo/cultural-consensus-musical-genres"
---

##

## Data
- We can treat the association data provided by respondents as **cognitive two-mode data**
  - In analogy with Krackhardt-style data for one-mode networks
  - Persons provide *perceived* associations between genres and social labels

## Average Perceptions
![](lizardo-skiles-tab3.png){width="85%"}

:::aside
Lizardo, O. and Skiles, S., 2016. Cultural objects as prisms: Perceived audience composition of musical genres as a resource for symbolic exclusion. *Socius*, 2, p.2378023116641695.
:::

## Stacked CA Analyses
- Three sets of scores:
  - row-scores are **person-specific** judgments of the relative similarity of genres and labels
  - column-scores represent the **aggregate judgment** of the relative similarity of genres (with respect to labels) and labels (with respect to genres)
  - Supplementary scores for genres and labels represent the **centroid** of the cloud of person-specific judgments

## Correspondence Plot of Genre Column Scores
![First and Second Dimensions](Plots/ca-genre-col12.png)

## Correspondence Plot of Genre Column Scores
![First and Third Dimensions](Plots/ca-genre-col13.png)

## Correspondence Plot of Genre Column Scores
![First and Fourth Dimensions](Plots/ca-genre-col14.png)

## Correspondence Plot of Label Column Scores
![First and Second Dimensions](Plots/ca-label-col12.png)

## Correspondence Plot of Label Column Scores
![First and Third Dimensions](Plots/ca-label-col13.png)

## Correspondence Plot of Label Column Scores
![First and Fourth Dimensions](Plots/ca-label-col14.png)

## Correspondence Plot of Genre Column Scores
![First and Second Dimensions](Plots/genre-sim-dens12.png)

## Correspondence Plot of Label Column Scores
![First and Second Dimensions](Plots/label-sim-dens12.png)

## Correspondence Plot of Genre Clouds
![First and Second Dimensions](Plots/genre-proto12.png)

## Correspondence Plot of Genre Clouds
![First and Third Dimensions](Plots/genre-proto13.png)

## Correspondence Plot of Genre Clouds
![First and Fourth Dimensions](Plots/genre-proto14.png)

## Correspondence Plot of Label Clouds
![First and Second Dimensions](Plots/label-proto12-1.png)

## Correspondence Plot of Label Clouds
![First and Second Dimensions](Plots/label-proto12-2.png)

## Correspondence Plot of Label Clouds
![First and Third Dimensions](Plots/label-proto13.png)

## Correspondence Plot of Label Clouds
![First and Fourth Dimensions](Plots/label-proto14.png)

## Genre Ridge Plot
![First Dimension](Plots/genre-ridge-dim1.png)

## Genre Ridge Plot
![Second Dimension](Plots/genre-ridge-dim2.png)

## Genre Ridge Plot
![Third Dimension](Plots/genre-ridge-dim3.png)

## Label Ridge Plot
![First Dimension](Plots/label-ridge-dim1.png)

## Label Ridge Plot
![Second Dimension](Plots/label-ridge-dim2.png)

## Label Ridge Plot
![Third Dimension](Plots/label-ridge-dim3.png)