### 2025-05-05 analysis-noproj.qmd
a buncha

---
title: "CA of Music Genre Stereotypes"
author: "Omar Lizardo"
date: "`r Sys.Date()`"
execute: 
  eval: true
  echo: true
  output: true
  warning: false
  message: false
---

```{r setup, include=FALSE}
   library(conflicted)
   library(dplyr)
   library(FactoMineR)
   library(ggplot2)
   library(ggrepel)
   library(corrplot)
   library(ggridges)
   library(haven)
   library(here)
   library(paletteer)
   cols <- paletteer_dynamic("cartography::multi.pal", 20)
   conflicts_prefer(dplyr::filter)
   conflicts_prefer(dplyr::select)
```

```{r import data and select genre stereotype variables}
   #importing data
   dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
   #replacing ids with character names
   dat <- bind_cols(tibble(id_char = paste("id", sapply(dat[, 1], as.character), sep = "_")), dat) %>% 
   select(id_char, classicalfemale:hvymtluc, -ends_with("none")) %>%  #selecting relevant columns
   na.omit() #removing missing values
   head(dat)
```

```{r Creating individual matrices}
      r <- c("Classical", "Opera", "Jazz", "Broadway", "Easy", "Big_Band", "Classic_Rock", "Country", "Bluegrass", "Folk", "Gospel", "Latin", "Rap", "Blues", "Reggae", "Pop", "Contemp_Rock", "AltRock", "Dance", "Metal")
       c <- c("Female", "Male", "White", "Black", "Hispanic", "Asian", "College", "No_College", "Young", "Mid_Age", "Old", "Lower_Class", "Working_Class", "Middle_Class", "Upper_Class")
   den <- rowSums(dat[,2:301])/300
   bad.den <- which(den > .95 | den < 0.05)
   dat2 <- dat[-c(bad.den), 2:301]
   ind.list <- split(dat2, seq(nrow(dat2))) # splitting each row of data frame into list objects
   ind.mat <- lapply(ind.list, function(x) {matrix(as.numeric(x), 20, 15, byrow = TRUE)}) # creating individual two mode network of genre by labels
   ind.mat.t <- lapply(ind.list, function(x) {matrix(as.numeric(x), 20, 15, byrow = TRUE)}) # creating individual two mode network of genre by labels
   ind.mat.t <- lapply(ind.mat.t, function(x) {t(x)})
   ind.mat <- lapply(ind.mat, function(x) {rownames(x) <- r; x}) # naming individual rows
   ind.mat <- lapply(ind.mat, function(x) {colnames(x) <- c; x}) # naming individual columns
   ind.mat.t <- lapply(ind.mat.t, function(x) {rownames(x) <- c; x}) # naming individual transposed rows
   ind.mat.t <- lapply(ind.mat.t, function(x) {colnames(x) <- r; x}) # naming individual transposed columns
   names(ind.mat) <- dat$id_char[-c(bad.den)]
   names(ind.mat.t) <- dat$id_char[-c(bad.den)]
   row.stack <- matrix(unlist(ind.mat), ncol = 15, byrow = TRUE) # genre stacked matrix
   col.stack <- matrix(unlist(ind.mat.t), ncol = 20, byrow = TRUE) # label stacked matrix
   colnames(row.stack) <- c # naming genre stacked matrix columns
   colnames(col.stack) <- r # naming label stacked matrix columns
   rownames(row.stack) <- rep(r, nrow(dat2)) # naming genre stacked matrix rows
   rownames(col.stack) <- rep(c, nrow(dat2)) # naming label stacked matrix rows
   zero.rows.r <- which(rowSums(row.stack) == 0) # which genre stacked matrix rows sum to zero?
   zero.rows.c <- which(rowSums(col.stack) == 0) # which label stacked matrix rows sum to zero?
```

```{r Data Examples (Two Mode)}
   png(file = here("Plots", "data-ex1.png"), width=700, height=700)
      corrplot(ind.mat[["id_65"]], is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 1, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1.75, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()
   png(file = here("Plots", "data-ex2.png"), width=700, height=700)
      corrplot(ind.mat[["id_236"]], is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 1, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1.75, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()
   png(file = here("Plots", "data-ex3.png"), width=700, height=700)
      corrplot(ind.mat[["id_1114"]], is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 1, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1.75, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()
   png(file = here("Plots", "data-ex4.png"), width=700, height=700)
      corrplot(ind.mat[["id_2002"]], is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 1, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1.75, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()

   ex.stack <- rbind(ind.mat[["id_65"]], ind.mat[["id_236"]])
   ex.stack.t <- rbind(t(ind.mat[["id_1114"]]), t(ind.mat[["id_2002"]]))
   png(file = here("Plots", "data-ex-stack.png"), width=700, height=700)
      corrplot(ex.stack, is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 0.75, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()
   png(file = here("Plots", "data-ex-stack-t.png"), width=700, height=700)
      corrplot(ex.stack.t, is.corr = FALSE, method = "color", 
         tl.col = "black", tl.cex = 0.75, tl.srt = 45,
         addCoef.col = "black", cl.pos = "n",
         number.cex = 1, number.digits = 0, addgrid.col = "darkgray",
         col=colorRampPalette(c("white","white","red"))(200))
   dev.off()
```

```{r CA}
   #genres
   row.stack.d <- data.frame(row.stack)
   row.stack.d$genre <- rep(r, nrow(dat2))
   row.stack.d$id <- rep(dat$id_char[-bad.den], each = 20)
   CA.r <- CA(row.stack.d, graph = FALSE, row.sup = zero.rows.r, quali.sup = c(16, 17))
   #labels
   col.stack.d <- data.frame(col.stack)
   col.stack.d$label <- rep(c, nrow(dat2))
   col.stack.d$id <- rep(dat$id_char[-bad.den], each = 15)
   CA.c <- CA(col.stack.d, graph = FALSE, row.sup = zero.rows.c, quali.sup = c(21, 22))
```

```{r Plot Data}
     make.plot.dat <- function(x) { 
         w <-  data.frame(x) %>% 
         mutate(col.obj = gsub("\\..*","",rownames(x))) 
         rownames(w) <- 1:nrow(x)
         return(w)
         }
      col.lab.dat <-  data.frame(CA.r$col$coord) #genre column scores
      col.gen.dat <-  data.frame(CA.c$col$coord) #label column scores
      sup.gen.dat <- CA.r$quali.sup$coord[1:20, ] 
      sup.lab.dat <- CA.c$quali.sup$coord[1:15, ] 
      sup.gen.dat <- data.frame(sup.gen.dat, 
                                col.obj = gsub("genre.", "", rownames(sup.gen.dat)))
      rownames(sup.gen.dat) <- sup.gen.dat$col.obj
      sup.lab.dat <- data.frame(sup.lab.dat, 
                                col.obj = gsub("label.", "", rownames(sup.lab.dat)))
      rownames(sup.lab.dat) <- sup.lab.dat$col.obj
      row.gen.dat <- make.plot.dat(CA.r$row$coord) #genre row scores
      row.lab.dat <- make.plot.dat(CA.c$row$coord) #label row scores
```

```{r Column and Supplementary Score Similarity Plots}
   h <- 625
   w <- 750
   source(here("make.sim.plot.R"))
   #genres
   png(file = here("Plots", "ca-genre-col12.png"), width=w, height=h)
   make.sim.plot(col.gen.dat, 1, 2, "red", l = 1)
   dev.off()
   png(file = here("Plots", "ca-genre-sup12.png"), width=w, height=h)
   make.sim.plot(sup.gen.dat, 1, 2, "red", l = .6)
   dev.off()
   # labels
   png(file = here("Plots", "ca-label-col12.png"), width=w, height=h)
   make.sim.plot(col.lab.dat, 1, 2, "blue", l = 1)
   dev.off()
   png(file = here("Plots", "ca-label-sup12.png"), width=w, height=h)
   make.sim.plot(sup.lab.dat, 1, 2, "blue", l = .6)
   dev.off()
```

```{r Genre Cloud Plots}
   source(here("make.fuzz.plot.R"))
   h <- 625
   w <- 750
   png(file = here("Plots", "genre-proto-ex1.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, sup.gen.dat, seed = 23456,
                     1, 2, xl = 2, yl = 2, pa = .05,
                     c("Folk", "Contemp_Rock", "Big_Band", "Classic_Rock"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex2.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, sup.gen.dat, seed = 234,
                     1, 2, xl = 2, yl = 2, pa = .05,
                     c("Reggae", "Opera", "Rap", "Opera"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex3.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, sup.gen.dat, seed = 234,
                     1, 2, xl = 2, yl = 1.5, pa = .05,
                     c("Gospel", "Latin", "Easy"))
   dev.off()
```

```{r Label Cloud Plots}
   source(here("make.fuzz.plot.R"))
   h <- 625
   w <- 750
   png(file = here("Plots", "label-proto-ex1.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, sup.lab.dat, seed = 234,
                     1, 2, xl = 2, yl = 2, pa = .05,
                     c("College", "Male", "Upper_Class"))
   dev.off()
   png(file = here("Plots", "label-proto-ex2.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, sup.lab.dat, seed = 234,
                     1, 2, xl = 2, yl = 2, pa = .05,
                     c("Female", "No_College", "White"))
   dev.off()
   png(file = here("Plots", "label-proto-ex3.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, sup.lab.dat, seed = 234,
                     1, 2, xl = 2, yl = 2.5, pa = .05,
                     c("Asian", "Working_Class", "Hispanic"))
   dev.off()
```

```{r Genre Ridge Plots}
      make.ridge.dat <- function(col.dat, ca.rows) {
        b <- col.dat %>% 
         rename(Prot.1 = Dim.1, Prot.2 = Dim.2, Prot.3 = Dim.3) %>% 
         mutate(col.obj = rownames(col.dat)) 
      x <-  data.frame(rbind(ca.rows)) %>% 
         mutate(col.obj = gsub("\\..*","",rownames(ca.rows))) %>% 
         select(-c("Dim.4", "Dim.5")) %>% 
         right_join(b) 
      return(x)
      }
      make.ridge.plot <- function(x, d1, d2, lim = 1) {
      p <- ggplot(data = x, aes(x = x[, d1], 
                                y = reorder(col.obj, x[, d2]),
                                group = col.obj, fill = col.obj, color = col.obj))
      p <- p + geom_density_ridges(alpha = .2, point_alpha = .2, jittered_points = TRUE)
      p <- p + guides(fill = "none", color = "none")
      p <- p + geom_vline(xintercept = 0, color = "black", linewidth = 1.25)
      p <- p + scale_color_manual(values = cols)
      p <- p + theme(axis.text = element_text(size = 20), 
                     axis.title = element_text(size = 24))
      p <- p + labs(x = paste("Dim.", d1))
      p <- p + theme_ridges() + xlim(-lim, lim) + labs(y = "") 
      return(p)
      }
   # genres
   ridge.dat.g <- make.ridge.dat(plot.dat1, CA.r$row$coord)
   png(file = here("Plots", "genre-ridge-dim1.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.g, 1, 5)
   dev.off()
   png(file = here("Plots", "genre-ridge-dim2.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.g, 2, 6)
   dev.off()
   png(file = here("Plots", "genre-ridge-dim3.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.g, 3, 7)
   dev.off()
   #labels
   ridge.dat.l <- make.ridge.dat(plot.dat2, CA.c$row$coord)
   png(file = here("Plots", "label-ridge-dim1.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.l, 1, 5)
   dev.off()
   png(file = here("Plots", "label-ridge-dim2.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.l, 2, 6)
   dev.off()
   png(file = here("Plots", "label-ridge-dim3.png"), width=w, height=h)
   make.ridge.plot(ridge.dat.l, 3, 7)
   dev.off()
```