### 2025-05-05 heidelberg-slides.qmd

---
title: "Musical Genres as Fuzzy Categories"
subtitle: ""
author: "Omar Lizardo"
institute: "*University of California, Los Angeles*"
date: "`r Sys.Date()`"
format:
  revealjs:
    incremental: false 
    theme: simple 
    smaller: true
    slide-number: true
    slide-level: 2
    progress: true
    footer: "https://github.com/olizardo/cultural-consensus-musical-genres"
---

##

## Data
- We can treat the association data provided by respondents as **cognitive two-mode data**
  - In analogy with Krackhardt-style data for one-mode networks
  - Persons provide *perceived* associations between genres and social labels

## Average Perceptions
![](lizardo-skiles-tab3.png){width="85%"}

:::aside
Lizardo, O. and Skiles, S., 2016. Cultural objects as prisms: Perceived audience composition of musical genres as a resource for symbolic exclusion. *Socius*, 2, p.2378023116641695.
:::

## Stacked CA Analyses
- Three sets of scores:
  - row-scores are **person-specific** judgments of the relative similarity of genres and labels
  - column-scores represent the **aggregate judgment** of the relative similarity of genres (with respect to labels) and labels (with respect to genres)
  - Supplementary scores for genres and labels represent the **centroid** of the cloud of person-specific judgments

## Mondo Breiger Approach

- Each individual provides a "personal" two mode network of genres by social labels with corresponding affiliation matrix $\mathbf{A}(p)$
  - $a^{(p)}_{gl} = 1$ if individual $p$ associates genre $g$ with label $l$
- The personal genre projection is given by:
  - $\mathbf{G}(p) = \mathbf{A}(p)\mathbf{A}(p)^T$
  - $g^{(p)}_{ij}$ records the $p^{th}$ individual's perceived similarity between genres $i$ and $j$ based on their shared labels
  - $g^{(p)}_{ii}$ records the number of times individual $p$ associates genre $i$ with a label

## Mondo Breiger Approach

- Each individual provides a "personal" two mode network of genres by social labels with corresponding affiliation matrix $\mathbf{A}(p)$
  - $a^{(p)}_{gl} = 1$ if individual $p$ associates genre $g$ with label $l$
- The personal label projection is given by:
  - $\mathbf{L}(p) = \mathbf{A}(p)^T\mathbf{A}(p)$
  - $l^{(p)}_{kl}$ records the $p^{th}$ individual's perceived similarity between labels $k$ and $l$ based on their shared genres
  - $l^{(p)}_{kk}$ records the number of times individual $p$ associates genre $i$ with a label

## Individual Matrix Examples (Two Mode)
:::: {.columns}
::: {.column width="50%"}
![id_65](Plots/data-ex1.png)
:::
::: {.column width="50%"}
![id_236](Plots/data-ex2.png)
:::
::::

## Individual Matrix Examples (Two Mode)
:::: {.columns}
::: {.column width="50%"}
![id_1114](Plots/data-ex3.png)
:::
::: {.column width="50%"}
![id_2002](Plots/data-ex4.png)
:::
::::

## Example Row Projection
:::: {.columns}
::: {.column width="50%"}
![id_65](Plots/data-ex1.png)
:::
::: {.column width="50%"}
![id_65](Plots/data-ex1-rp.png)
:::
::::

##  Example Column Projection
:::: {.columns}
::: {.column width="50%"}
![id_65](Plots/data-ex1.png)
:::
::: {.column width="50%"}
![id_65](Plots/data-ex1-cp.png)
:::
::::

##  Example Row Projection
:::: {.columns}
::: {.column width="50%"}
![id_236](Plots/data-ex2.png)
:::
::: {.column width="50%"}
![id_236](Plots/data-ex2-rp.png)
:::
::::

##  Example Column Projection
:::: {.columns}
::: {.column width="50%"}
![id_65](Plots/data-ex2.png)
:::
::: {.column width="50%"}
![id_65](Plots/data-ex2-cp.png)
:::
::::

##  Stacked Matrices Examples
:::: {.columns}
::: {.column width="50%"}
![id_65 and id_236 (Rows)](Plots/data-ex-stack1.png)
:::
::: {.column width="50%"}
![id_65 and id_236 (Columns)](Plots/data-ex-stack2.png)
:::
::::

## Correspondence Plot of Label Column Scores
![First and Second Dimensions](Plots/ca-label-col12.png)

## Correspondence Plot of Label Column Scores
![First and Third Dimensions](Plots/ca-label-col13.png)

## Correspondence Plot of Genre Column Scores
![First and Second Dimensions](Plots/ca-genre-col12.png)

## Correspondence Plot of Genre Column Scores
![First and Third Dimensions](Plots/ca-genre-col13.png)

## Correspondence Plot of Genre Clouds
![First and Second Dimensions](Plots/genre-proto-ex1.png)

## Correspondence Plot of Genre Clouds
![First and Second Dimensions](Plots/genre-proto-ex2.png)

## Correspondence Plot of Genre Clouds
![First and Second Dimensions](Plots/genre-proto-ex3.png)

## Correspondence Plot of Label Clouds
![First and Second Dimensions](Plots/label-proto-ex1.png)

## Correspondence Plot of Label Clouds
![First and Second Dimensions](Plots/label-proto-ex2.png)

## Correspondence Plot of Label Clouds
![First and Second Dimensions](Plots/label-proto-ex3.png)

## Correspondence Plot of Label Clouds
![First and Second Dimensions](Plots/label-proto-ex4.png)